Psychological Needs During Intergroup Contact

### 2022-01-06 Methods-and-Results.Rmd
Feedback Max M&R

\subsubsection{Contact Hypothesis}We tested the most general contact hypothesis in two steps. First, we assessed whether more intergroup interactions were related to to more positive outgroup attitudes. Second, we tested whether a potential positive effect on outgroup attitudes depended on the interaction quality (jointly with the number of interactions). We find that neither the number of interactions nor the number of daily diary responses with an interaction were significantly related with the average outgroup attitudes (all |\textit{r}| < `r rContactAvgMax`, all \textit{p} > `r pContactAvgMin`). This is to say that within our data, participants with more outgroup interactions did not have significantly more positive outgroup attitudes. This might be due to the aggregation within the participants or the small sample size of between participant data. Nonetheless, the aggregated data does not support the notion that simply having more interactions with an outgroup results in more positive outgroup attitudes.

@@

However, despite the missing relationship with the number of interactions, we find a medium sized correlation between the participants' Average Interaction Quality and their Average Outgroup Attitudes. Thus within our data participants with a higher quality outgroup interactions also held more positive attitudes towards that group. And when considering the number of interactions and average interaction quality jointly in a linear regression, we additionally find a statistically significant interaction term (\textit{b} = `r coef(summary(mdlWorker$lmAttFreqQualX))["SumContactNL_c:AvQuality_c","Estimate"] %>% round(2) %>% format(nsmall=2)`, \textit{t}(`r mdlWorker$lmAttFreqQualX$df.residual`) = `r coef(summary(mdlWorker$lmAttFreqQualX))["SumContactNL_c:AvQuality_c","t value"] %>% round(2) %>% format(nsmall=2)`, \textit{p} = `r coef(summary(mdlWorker$lmAttFreqQualX))["SumContactNL_c:AvQuality_c","Pr(>|t|)"] %>% round(3) %>% format(nsmall=3)`, $\eta_p^2$ = `r mdlWorker$lmAttFreqQualXEta %>% filter(Parameter == "SumContactNL_c:AvQuality_c") %>% select(Eta2_partial) %>% as.numeric %>% round(2) %>% format(nsmall=2)`). Looking at a floodlight analysis of the effect, we find that in our sample with an increasing number of interactions the positive effect of average interaction quality becomes weaker. However, it should be noted that this is based on data aggregating all within participant nuances and is only the date of 21 people.

@@

We additionally used a multilevel regression to check whether having an interaction with an outgroup member had a situational (i.e., contemporaneous) effect within the participants. We find that having an outgroup interaction is indeed associated with significantly more positive outgroup attitudes within the participants (`r MlCoeffLatex(lmeMdl = mdlWorker$lmeInterceptAttType, lmerCI = mdlWorker$lmerInterceptAttTypeCI, varName = "OutgroupInteractionYes")`), even after controlling for having an interaction with a non-Dutch (which did not relate to outgroup attitudes independently; For full results see Online Supplementary Materials B)\footnote{Interestingly, adding random slopes to this model did not explain additional variance. This is unusual and might indicate the the effect is very consistent across participants. However, the small number of participants, or other measurement issues provide an alternative explanation}. Thus, in our first data we find mixed results, where outgroup contacts show a positive effect on outgroup attitudes in the moment but these results do not translate into average differences between participants, keeping in mind that the between-participant analyses in the aggregated form are based in a small sample size.

\subsubsection{Core Need}The main proposal of our article is that the success of an outgroup interaction might be explained by whether or not the interaction fulfilled the person's core situational need. This should, in turn, be due to a higher perceived interaction quality. We will sequentially test whether the fulfillment of the core need during an interaction is (1) related to more positive outgroup attitudes, (2) higher perceived interaction quality, and (3) whether the variance explained by the core need is assumed by the perceived interaction quality if considered jointly.

@@

We find that in the highest multilevel models, the fulfillment of core situational needs during outgroup contacts was associated with more positive outgroup attitudes (random slopes model; `r MlCoeffLatex(lmeMdl = mdlWorkerOut$Att$lmeSlopesCore, lmerCI = mdlWorkerOut$Att$lmerSlopesCoreCI, varName = "keymotive_fulfillemt_1_cwc")`) and also predicted higher perceived interaction quality (random intercept model; `r MlCoeffLatex(lmeMdl = mdlWorkerOut$Qlt$lmeInterceptCore, lmerCI = mdlWorkerOut$Qlt$lmerInterceptCoreCI, varName = "keymotive_fulfillemt_1_cwc")`). Moreover, when we consider the influences of core need fulfillment and interaction quality on outgroup attitudes jointly, we find that virtually all variance is explained by perceived interaction quality (random slopes mode; `r MlCoeffLatex(lmeMdl = mdlWorkerOut$Att$lmeSlopesCoreQlt, lmerCI = mdlWorkerOut$Att$lmerSlopesCoreQltCI, varName = "quality_overall_1_cwc")`) and only little unique variance is still explained by core need fulfillment (`r MlCoeffLatex(lmeMdl = mdlWorkerOut$Att$lmeSlopesCoreQlt, lmerCI = mdlWorkerOut$Att$lmerSlopesCoreQltCI, varName = "keymotive_fulfillemt_1_cwc")`). We thus find support for our hypotheses and can conclude that in this data set the fulfillment of core situational needs had a significant influence on outgroup attitudes; this effect likely being explained by its effect through perceived interaction quality.

@@

To build further confidence in our results, we assessed two additional models that might offer alternative explanations. First, to ensure that the effect of core need fulfillment is specific to an actual interaction, we compared it core need fulfillement in situations without an interaction. For this, we analyzed the generalized situational core need fulfillment (either during an interaction or about the daytime in general) and tested whether the effect differed during daily diary measurements with and without outgroup contacts. We found no main effect of core need fulfillment (random slopes model; `r MlCoeffLatex(lmeMdl = mdlWorker$lmeSlopesAttCoreInt, lmerCI = mdlWorker$lmerSlopesAttCoreIntCI, varName = "keyMotiveFulfilled_cwc")`) but a significant interaction effect of core need fulfillment and outgroup contact (`r MlCoeffLatex(lmeMdl = mdlWorker$lmeSlopesAttCoreInt, lmerCI = mdlWorker$lmerSlopesAttCoreIntCI, varName = "keyMotiveFulfilled_cwc:OutgroupInteraction")`). Together with a significant main effect of having an outgroup contact, this indicates that it is not key need fulfillment in general --- but only key need fulfillment during an outgroup contact that predicts more positive outgroup attitudes.

@@

The aim of Study 2 is similar to Study 1, as we again test the general contact hypothesis, the influence of core need fulfillment, and perceived interaction quality during intergroup contacts. However, in this second study we collected a substantially larger sample of international students who recently arrived in the Netherlands and also improved the study design (e.g. pop up explanations described later). The survey method again offers a large body of ecologically valid data on need satisfaction in real-life intergroup contact situations as these students will likely interact with the Dutch majority outgroup on a daily basis. Data was collected from November 19th, 2018 through January 6th, 2019.

The full surveys are available in our Online Supplementary Material A and the full data description is available in Online Supplementary Materials B. Correlations and descriptive statistics of the included variables are available in Table \ref{tab:studentVarDescr} and Table \ref{tab:studentOutVarDescr}.

@@

\paragraph{Intergroup Contact}To measure intergroup contacts, every experience sampling measurement started with the question “\textit{Did you meet a Dutch person this morning [/afternoon]? (in-person interaction for at least 10 minutes)}”. Participants were additionally offered a pop-up explanation: "With in-person interaction, we mean a continued interaction with another person (potentially in a group) that lasted at least 10 minutes. This interaction should be offline and face-to-face. It should include some form of verbal communication and should be uninterrupted to still count as the same interaction. Any individual interaction can last minutes or hours. If there were multiple interaction partners, we would like you to focus on the person that was most important to you during the interaction.". The participants recorded between `r dtStudentSupp$studentContactFreq$SumContactNL[dtStudentSupp$studentContactFreq$SumContactNL > 0] %>% min`--`r dtStudentSupp$studentContactFreq$SumContactNL %>% max`
(`r dtStudentSupp$studentContactFreq$PercContactNL[dtStudentSupp$studentContactFreq$PercContactNL > 0] %>% min %>% round(2) %>% format(nsmall=2)`--`r dtStudentSupp$studentContactFreq$PercContactNL %>% max %>% round(2) %>% format(nsmall=2)`\% of individual daily diary measurements; `r (sum(dtStudentSupp$studentContactFreq$SumContactNL) / nrow(dtStudentSupp$studentInteractionType) * 100) %>% round(2) %>% format(nsmall=2)`\% of all `r nrow(dtStudentSupp$studentInteractionType)` daily diary responses).

@@

For this study we specifically recruited international medical students, because they represent a particular group of migrants who face structural requirements to integrate and interact with Dutch majority outgroup members on a daily basis. As part of their educational program, the migrants are required to take language courses and interact with patients as part of their medical internships and medical residency. The extensive longitudinal survey method again offers a large body of ecologically valid data on need satisfaction in real-life intergroup contact situations. Data were collected from November 8th, 2019 to January 10th, 2020.

The full surveys are available in our Online Supplementary Material A and the full data description is available in Online Supplementary Materials B. Correlations and descriptive statistics of the included variables are available in Table \ref{tab:medicalVarDescr} and Table \ref{tab:medicalOutVarDescr}.

>>>>

\subsubsection{Contact Hypothesis}We tested the most general contact hypothesis in two steps. First, we assessed whether more intergroup interactions were related to to more positive outgroup attitudes. Second, we tested whether a potential positive effect on outgroup attitudes depended on the interaction quality (jointly with the number of interactions). We find that neither the number of interactions nor the number of daily diary responses with an interaction were significantly related with the average outgroup attitudes (all | \textit{r} | < `r rContactAvgMax`, all \textit{p} > `r pContactAvgMin`). This is to say that within our data, participants with more outgroup interactions did not have significantly more positive outgroup attitudes. This might be due to the aggregation within the participants or the small sample size of between participant data. Nonetheless, the aggregated data does not support the notion that simply having more interactions with an outgroup results in more positive outgroup attitudes.

@@

However, despite the missing relationship with the number of interactions, we find a medium sized correlation between the participants' Average Interaction Quality and their Average Outgroup Attitudes. Thus within our data participants with a higher quality outgroup interactions also held more positive attitudes towards that group. And when considering the number of interactions and average interaction quality jointly in a linear regression, we additionally find a statistically significant interaction term (\textit{b} = `r coef(summary(mdlWorker$lmAttFreqQualX))["SumContactNL_c:AvQuality_c","Estimate"] %>% round(2) %>% format(nsmall=2)`, \textit{t}(`r mdlWorker$lmAttFreqQualX$df.residual`) = `r coef(summary(mdlWorker$lmAttFreqQualX))["SumContactNL_c:AvQuality_c","t value"] %>% round(2) %>% format(nsmall=2)`, \textit{p} = `r coef(summary(mdlWorker$lmAttFreqQualX))["SumContactNL_c:AvQuality_c","Pr(>|t|)"] %>% round(3) %>% format(nsmall=3)`, $\eta_p^2$ = `r mdlWorker$lmAttFreqQualXEta %>% filter(Parameter == "SumContactNL_c:AvQuality_c") %>% select(Eta2_partial) %>% as.numeric %>% round(2) %>% format(nsmall=2)`). Looking at a floodlight analysis of the effect, we find that in our sample with an increasing number of interactions the positive effect of average interaction quality becomes weaker. However, it should be noted that this is based on data aggregating all within participant nuances and is only based on 21 people.

@@

We additionally used a multilevel regression to check whether having an interaction with an outgroup member had a situational (i.e., contemporaneous) effect within the participants. We find that having an outgroup interaction is indeed associated with significantly more positive outgroup attitudes within the participants (`r MlCoeffLatex(lmeMdl = mdlWorker$lmeInterceptAttType, lmerCI = mdlWorker$lmerInterceptAttTypeCI, varName = "OutgroupInteractionYes")`), even after controlling for having an interaction with a non-Dutch (which did not relate to outgroup attitudes independently; For full results see Online Supplementary Materials B)\footnote{Interestingly, adding random slopes to this model did not explain additional variance. This is unusual and might indicate the the effect is very consistent across participants. However, the small number of participants, or other measurement issues provide an alternative explanation}. Thus, in our first data we find mixed results, where outgroup contacts show a positive effect on outgroup attitudes in the moment but these results do not translate into average differences between participants, keeping in mind that the between-participant analyses in the aggregated form are based on a small sample size.

\subsubsection{Core Need}The main proposal of our article is that the success of an outgroup interaction might be explained by whether or not the interaction fulfilled the person's core situational need. This should, in turn, be due to a higher perceived interaction quality. We sequentially test whether the fulfillment of the core need during an interaction is (1) related to more positive outgroup attitudes, (2) higher perceived interaction quality, and (3) whether the variance explained by the core need is assumed by the perceived interaction quality if considered jointly.

@@

We find that in the highest multilevel models, the fulfillment of core situational needs during outgroup contacts was associated with more positive outgroup attitudes (random slopes model; `r MlCoeffLatex(lmeMdl = mdlWorkerOut$Att$lmeSlopesCore, lmerCI = mdlWorkerOut$Att$lmerSlopesCoreCI, varName = "keymotive_fulfillemt_1_cwc")`) and also predicted higher perceived interaction quality (random intercept model; `r MlCoeffLatex(lmeMdl = mdlWorkerOut$Qlt$lmeInterceptCore, lmerCI = mdlWorkerOut$Qlt$lmerInterceptCoreCI, varName = "keymotive_fulfillemt_1_cwc")`). Moreover, when we consider the influences of core need fulfillment and interaction quality on outgroup attitudes jointly, we find that virtually all variance is explained by perceived interaction quality (random slopes mode; `r MlCoeffLatex(lmeMdl = mdlWorkerOut$Att$lmeSlopesCoreQlt, lmerCI = mdlWorkerOut$Att$lmerSlopesCoreQltCI, varName = "quality_overall_1_cwc")`) and only little unique variance is still explained by core need fulfillment (`r MlCoeffLatex(lmeMdl = mdlWorkerOut$Att$lmeSlopesCoreQlt, lmerCI = mdlWorkerOut$Att$lmerSlopesCoreQltCI, varName = "keymotive_fulfillemt_1_cwc")`). We thus find support for our hypotheses and can conclude that in this data set the fulfillment of core situational needs had a significant influence on outgroup attitudes. Additionally, this effect is likely being explained by its effect through perceived interaction quality.

@@

To build further confidence in our results, we assessed two additional models that might offer alternative explanations. First, to ensure that the effect of core need fulfillment is specific to an actual interaction, we compared the effect to core need fulfillement in situations without an interaction. For this, we analyzed the generalized situational core need fulfillment (either during an interaction or about the daytime in general) and tested whether the effect differed during daily diary measurements with and without outgroup contacts. We found no main effect of core need fulfillment (random slopes model; `r MlCoeffLatex(lmeMdl = mdlWorker$lmeSlopesAttCoreInt, lmerCI = mdlWorker$lmerSlopesAttCoreIntCI, varName = "keyMotiveFulfilled_cwc")`) but a significant interaction effect of core need fulfillment and outgroup contact (`r MlCoeffLatex(lmeMdl = mdlWorker$lmeSlopesAttCoreInt, lmerCI = mdlWorker$lmerSlopesAttCoreIntCI, varName = "keyMotiveFulfilled_cwc:OutgroupInteraction")`). Together with a significant main effect of having an outgroup contact, this indicates that it is not key need fulfillment in general --- but only key need fulfillment during an outgroup contact that predicts more positive outgroup attitudes.

@@

The aim of Study 2 is similar to Study 1, as we again test the general contact hypothesis, the influence of core need fulfillment, and perceived interaction quality during intergroup contacts. However, in this second study we collected a substantially larger sample of international students who recently arrived in the Netherlands and also improved the study design (e.g., pop up explanations described later). The survey method again offers a large body of ecologically valid data on need satisfaction in real-life intergroup contact situations as these students will likely interact with the Dutch majority outgroup on a daily basis. Data was collected from November 19\textsuperscript{th}, 2018 through January 6\textsuperscript{th}, 2019.

@@

\paragraph{Intergroup Contact}To measure intergroup contacts, every experience sampling measurement started with the question “\textit{Did you meet a Dutch person this morning [/afternoon]? (in-person interaction for at least 10 minutes)}”. Participants were additionally offered a pop-up explanation: "With in-person interaction, we mean a continued interaction with another person (potentially in a group) that lasted at least 10 minutes. This interaction should be offline and face-to-face. It should include some form of verbal communication and should be uninterrupted to still count as the same interaction. Any individual interaction can last minutes or hours. If there were multiple interaction partners, we would like you to focus on the person that was most important to you during the interaction.". The participants recorded between `r dtStudentSupp$studentContactFreq$SumContactNL[dtStudentSupp$studentContactFreq$SumContactNL > 0] %>% min`--`r dtStudentSupp$studentContactFreq$SumContactNL %>% max`
interactions with Dutch majority people (`r dtStudentSupp$studentContactFreq$PercContactNL[dtStudentSupp$studentContactFreq$PercContactNL > 0] %>% min %>% round(2) %>% format(nsmall=2)`--`r dtStudentSupp$studentContactFreq$PercContactNL %>% max %>% round(2) %>% format(nsmall=2)`\% of individual daily diary measurements; `r (sum(dtStudentSupp$studentContactFreq$SumContactNL) / nrow(dtStudentSupp$studentInteractionType) * 100) %>% round(2) %>% format(nsmall=2)`\% of all `r nrow(dtStudentSupp$studentInteractionType)` daily diary responses).

@@

For this study we specifically recruited international medical students, because they represent a particular group of migrants who face structural requirements to integrate and interact with Dutch majority outgroup members on a daily basis. As part of their educational program, the migrants are required to take language courses and interact with patients as part of their medical internships and medical residency. The extensive longitudinal survey method again offers a large body of ecologically valid data on need satisfaction in real-life intergroup contact situations. Data were collected from November 8\textsuperscript{th}, 2019 to January 10\textsuperscript{th}, 2020.

The full surveys are available in our Online Supplementary Material A and the full data description is available in Online Supplementary Materials B. Correlations and descriptive statistics of the included variables are available in Tables \ref{tab:descrFullLong} and \ref{tab:descrOutLong}.

### 2022-01-06 Methods-and-Results.tex
Feedback Max M&R

We tested the most general contact hypothesis in two steps. First, we
assessed whether more intergroup interactions were related to to more
positive outgroup attitudes. Second, we tested whether a potential
positive effect on outgroup attitudes depended on the interaction
quality (jointly with the number of interactions). We find that neither
the number of interactions nor the number of daily diary responses with
an interaction were significantly related with the average outgroup
attitudes (all \textbar{}\textit{r}\textbar{} \textless{} 0.03, all
\textit{p} \textgreater{} 0.887). This is to say that within our data,
participants with more outgroup interactions did not have significantly
more positive outgroup attitudes. This might be due to the aggregation
within the participants or the small sample size of between participant
data. Nonetheless, the aggregated data does not support the notion that
simply having more interactions with an outgroup results in more
positive outgroup attitudes.

However, despite the missing relationship with the number of
interactions, we find a medium sized correlation between the
participants' Average Interaction Quality and their Average Outgroup
Attitudes. Thus within our data participants with a higher quality
outgroup interactions also held more positive attitudes towards that
group. And when considering the number of interactions and average
interaction quality jointly in a linear regression, we additionally find
a statistically significant interaction term (\textit{b} = -0.05,
\textit{t}(17) = -2.95, \textit{p} = 0.009, \(\eta_p^2\) = 0.34).
Looking at a floodlight analysis of the effect, we find that in our
sample with an increasing number of interactions the positive effect of
average interaction quality becomes weaker. However, it should be noted
that this is based on data aggregating all within participant nuances
and is only the date of 21 people.

We additionally used a multilevel regression to check whether having an
interaction with an outgroup member had a situational (i.e.,
contemporaneous) effect within the participants. We find that having an
outgroup interaction is indeed associated with significantly more
positive outgroup attitudes within the participants (\textit{b} = 2.48,
\textit{t}(1200) = 4.36, \textit{p} \textless{} .001,
\textit{95\%CI}{[}1.36, 3.59{]}), even after controlling for having an
interaction with a non-Dutch (which did not relate to outgroup attitudes
independently; For full results see Online Supplementary Materials
B)\footnote{Interestingly, adding random slopes to this model did not explain additional variance. This is unusual and might indicate the the effect is very consistent across participants. However, the small number of participants, or other measurement issues provide an alternative explanation}.
Thus, in our first data we find mixed results, where outgroup contacts
show a positive effect on outgroup attitudes in the moment but these
results do not translate into average differences between participants,
keeping in mind that the between-participant analyses in the aggregated
form are based in a small sample size.

@@

The main proposal of our article is that the success of an outgroup
interaction might be explained by whether or not the interaction
fulfilled the person's core situational need. This should, in turn, be
due to a higher perceived interaction quality. We will sequentially test
whether the fulfillment of the core need during an interaction is (1)
related to more positive outgroup attitudes, (2) higher perceived
interaction quality, and (3) whether the variance explained by the core
need is assumed by the perceived interaction quality if considered
jointly.

We find that in the highest multilevel models, the fulfillment of core
situational needs during outgroup contacts was associated with more
positive outgroup attitudes (random slopes model; \textit{b} = 0.17,
\textit{t}(365) = 2.93, \textit{p} = 0.004, \textit{95\%CI}{[}0.06,
0.29{]}) and also predicted higher perceived interaction quality (random
intercept model; \textit{b} = 0.42, \textit{t}(365) = 8.07, \textit{p}
\textless{} .001, \textit{95\%CI}{[}0.32, 0.52{]}). Moreover, when we
consider the influences of core need fulfillment and interaction quality
on outgroup attitudes jointly, we find that virtually all variance is
explained by perceived interaction quality (random slopes mode;
\textit{b} = 0.23, \textit{t}(364) = 4.22, \textit{p} \textless{} .001,
\textit{95\%CI}{[}0.12, 0.33{]}) and only little unique variance is
still explained by core need fulfillment (\textit{b} = 0.04,
\textit{t}(364) = 0.64, \textit{p} = 0.523, \textit{95\%CI}{[}-0.08,
0.15{]}). We thus find support for our hypotheses and can conclude that
in this data set the fulfillment of core situational needs had a
significant influence on outgroup attitudes; this effect likely being
explained by its effect through perceived interaction quality.

@@

To build further confidence in our results, we assessed two additional
models that might offer alternative explanations. First, to ensure that
the effect of core need fulfillment is specific to an actual
interaction, we compared it core need fulfillement in situations without
an interaction. For this, we analyzed the generalized situational core
need fulfillment (either during an interaction or about the daytime in
general) and tested whether the effect differed during daily diary
measurements with and without outgroup contacts. We found no main effect
of core need fulfillment (random slopes model; \textit{b} = -0.10,
\textit{t}(1199) = -2.22, \textit{p} = 0.026, \textit{95\%CI}{[}-0.18,
-0.01{]}) but a significant interaction effect of core need fulfillment
and outgroup contact (\textit{b} = 0.13, \textit{t}(1199) = 4.31,
\textit{p} \textless{} .001, \textit{95\%CI}{[}0.07, 0.18{]}). Together
with a significant main effect of having an outgroup contact, this
indicates that it is not key need fulfillment in general --- but only
key need fulfillment during an outgroup contact that predicts more
positive outgroup attitudes.

@@

The aim of Study 2 is similar to Study 1, as we again test the general
contact hypothesis, the influence of core need fulfillment, and
perceived interaction quality during intergroup contacts. However, in
this second study we collected a substantially larger sample of
international students who recently arrived in the Netherlands and also
improved the study design (e.g.~pop up explanations described later).
The survey method again offers a large body of ecologically valid data
on need satisfaction in real-life intergroup contact situations as these
students will likely interact with the Dutch majority outgroup on a
daily basis. Data was collected from November 19th, 2018 through January
6th, 2019.

The full surveys are available in our Online Supplementary Material A
and the full data description is available in Online Supplementary
Materials B. Correlations and descriptive statistics of the included
variables are available in Table \ref{tab:studentVarDescr} and Table
\ref{tab:studentOutVarDescr}.

@@

To measure intergroup contacts, every experience sampling measurement
started with the question
``\textit{Did you meet a Dutch person this morning [/afternoon]? (in-person interaction for at least 10 minutes)}''.
Participants were additionally offered a pop-up explanation: ``With
in-person interaction, we mean a continued interaction with another
person (potentially in a group) that lasted at least 10 minutes. This
interaction should be offline and face-to-face. It should include some
form of verbal communication and should be uninterrupted to still count
as the same interaction. Any individual interaction can last minutes or
hours. If there were multiple interaction partners, we would like you to
focus on the person that was most important to you during the
interaction.''. The participants recorded between 1--43 (1.64--93.75\%
of individual daily diary measurements; 18.83\% of all 4965 daily diary
responses).

>>>>

We tested the most general contact hypothesis in two steps. First, we
assessed whether more intergroup interactions were related to to more
positive outgroup attitudes. Second, we tested whether a potential
positive effect on outgroup attitudes depended on the interaction
quality (jointly with the number of interactions). We find that neither
the number of interactions nor the number of daily diary responses with
an interaction were significantly related with the average outgroup
attitudes (all \textbar{} \textit{r} \textbar{} \textless{} 0.03, all
\textit{p} \textgreater{} 0.887). This is to say that within our data,
participants with more outgroup interactions did not have significantly
more positive outgroup attitudes. This might be due to the aggregation
within the participants or the small sample size of between participant
data. Nonetheless, the aggregated data does not support the notion that
simply having more interactions with an outgroup results in more
positive outgroup attitudes.

However, despite the missing relationship with the number of
interactions, we find a medium sized correlation between the
participants' Average Interaction Quality and their Average Outgroup
Attitudes. Thus within our data participants with a higher quality
outgroup interactions also held more positive attitudes towards that
group. And when considering the number of interactions and average
interaction quality jointly in a linear regression, we additionally find
a statistically significant interaction term (\textit{b} = -0.05,
\textit{t}(17) = -2.95, \textit{p} = 0.009, \(\eta_p^2\) = 0.34).
Looking at a floodlight analysis of the effect, we find that in our
sample with an increasing number of interactions the positive effect of
average interaction quality becomes weaker. However, it should be noted
that this is based on data aggregating all within participant nuances
and is only based on 21 people.

We additionally used a multilevel regression to check whether having an
interaction with an outgroup member had a situational (i.e.,
contemporaneous) effect within the participants. We find that having an
outgroup interaction is indeed associated with significantly more
positive outgroup attitudes within the participants (\textit{b} = 2.48,
\textit{t}(1200) = 4.36, \textit{p} \textless{} .001,
\textit{95\%CI}{[}1.36, 3.59{]}), even after controlling for having an
interaction with a non-Dutch (which did not relate to outgroup attitudes
independently; For full results see Online Supplementary Materials
B)\footnote{Interestingly, adding random slopes to this model did not explain additional variance. This is unusual and might indicate the the effect is very consistent across participants. However, the small number of participants, or other measurement issues provide an alternative explanation}.
Thus, in our first data we find mixed results, where outgroup contacts
show a positive effect on outgroup attitudes in the moment but these
results do not translate into average differences between participants,
keeping in mind that the between-participant analyses in the aggregated
form are based on a small sample size.

@@

The main proposal of our article is that the success of an outgroup
interaction might be explained by whether or not the interaction
fulfilled the person's core situational need. This should, in turn, be
due to a higher perceived interaction quality. We sequentially test
whether the fulfillment of the core need during an interaction is (1)
related to more positive outgroup attitudes, (2) higher perceived
interaction quality, and (3) whether the variance explained by the core
need is assumed by the perceived interaction quality if considered
jointly.

We find that in the highest multilevel models, the fulfillment of core
situational needs during outgroup contacts was associated with more
positive outgroup attitudes (random slopes model; \textit{b} = 0.17,
\textit{t}(365) = 2.93, \textit{p} = 0.004, \textit{95\%CI}{[}0.06,
0.29{]}) and also predicted higher perceived interaction quality (random
intercept model; \textit{b} = 0.42, \textit{t}(365) = 8.07, \textit{p}
\textless{} .001, \textit{95\%CI}{[}0.32, 0.52{]}). Moreover, when we
consider the influences of core need fulfillment and interaction quality
on outgroup attitudes jointly, we find that virtually all variance is
explained by perceived interaction quality (random slopes mode;
\textit{b} = 0.23, \textit{t}(364) = 4.22, \textit{p} \textless{} .001,
\textit{95\%CI}{[}0.12, 0.33{]}) and only little unique variance is
still explained by core need fulfillment (\textit{b} = 0.04,
\textit{t}(364) = 0.64, \textit{p} = 0.523, \textit{95\%CI}{[}-0.08,
0.15{]}). We thus find support for our hypotheses and can conclude that
in this data set the fulfillment of core situational needs had a
significant influence on outgroup attitudes. Additionally, this effect
is likely being explained by its effect through perceived interaction
quality.

@@

To build further confidence in our results, we assessed two additional
models that might offer alternative explanations. First, to ensure that
the effect of core need fulfillment is specific to an actual
interaction, we compared the effect to core need fulfillement in
situations without an interaction. For this, we analyzed the generalized
situational core need fulfillment (either during an interaction or about
the daytime in general) and tested whether the effect differed during
daily diary measurements with and without outgroup contacts. We found no
main effect of core need fulfillment (random slopes model; \textit{b} =
-0.10, \textit{t}(1199) = -2.22, \textit{p} = 0.026,
\textit{95\%CI}{[}-0.18, -0.01{]}) but a significant interaction effect
of core need fulfillment and outgroup contact (\textit{b} = 0.13,
\textit{t}(1199) = 4.31, \textit{p} \textless{} .001,
\textit{95\%CI}{[}0.07, 0.18{]}). Together with a significant main
effect of having an outgroup contact, this indicates that it is not key
need fulfillment in general --- but only key need fulfillment during an
outgroup contact that predicts more positive outgroup attitudes.

@@

The aim of Study 2 is similar to Study 1, as we again test the general
contact hypothesis, the influence of core need fulfillment, and
perceived interaction quality during intergroup contacts. However, in
this second study we collected a substantially larger sample of
international students who recently arrived in the Netherlands and also
improved the study design (e.g., pop up explanations described later).
The survey method again offers a large body of ecologically valid data
on need satisfaction in real-life intergroup contact situations as these
students will likely interact with the Dutch majority outgroup on a
daily basis. Data was collected from November 19\textsuperscript{th},
2018 through January 6\textsuperscript{th}, 2019.

@@

To measure intergroup contacts, every experience sampling measurement
started with the question
``\textit{Did you meet a Dutch person this morning [/afternoon]? (in-person interaction for at least 10 minutes)}''.
Participants were additionally offered a pop-up explanation: ``With
in-person interaction, we mean a continued interaction with another
person (potentially in a group) that lasted at least 10 minutes. This
interaction should be offline and face-to-face. It should include some
form of verbal communication and should be uninterrupted to still count
as the same interaction. Any individual interaction can last minutes or
hours. If there were multiple interaction partners, we would like you to
focus on the person that was most important to you during the
interaction.''. The participants recorded between 1--43 interactions
with Dutch majority people (1.64--93.75\% of individual daily diary
measurements; 18.83\% of all 4965 daily diary responses).
